Type Error Telemetry at Scale: Pseudonymized and Stochiastic

### 2023-09-19 programming23/paper.tex
quick half-draft report on specific type errors

@@

%% \paragraph{Most Type Errors are Typos}
One striking aspect of~\cref{t:type-error-survival} is the large number
of \code{SyntaxError}s and the related errors \code{UnknownSymbol} and
\code{UnknownProperty}.
Most type errors in the data are likely due to typos.
\Cref{s:type-error-count} discusses this point in detail.
For now, we remark that \code{SyntaxError} is the only possible error
in \mnocheck{} mode (by design) and observe that these errors are
usually in the Drop column and seldom in the Keep column.

A second observation is that there are relatively few errors that overlap
with the edit range; the numbers are low in general.
The highest value in the \mstrict{} columns is 71, meaning 71 records
drop a syntax error.
This number is very small relative to the total number of \mstrict{} records
from~\cref{f:dataset-overview} (6,505), and similarly for
the \mnonstrict{} and \mnocheck{} modes.
%% FILL show percentages for all three modes?
Edits seldom overlap with type errors.
\FILL{} what does this mean? Not sure if bad, or inconclusive.

When edits do overlap, however, ... good?

@@

@@

%% use the ... percent of all records??
%% total 1.3M nocheck, 156K nonstrict, 6K strict
%% ... uhhh wow so few errors!

\begin{table}[t]
  \caption{Type error survival}
  \label{t:type-error-survival}

    \begin{tabular}{lr@{}r@{}rr@{}r@{}rr@{}r@{}r}
      & \zerowidth{\mnocheck{}} & & & \zerowidth{\mnonstrict{}} & & & \zerowidth{\mstrict{}} & & \\
      & \rbox{Add} & \ybox{Keep} & \gbox{Drop} & \rbox{Add} & \ybox{Keep} & \gbox{Drop} & \rbox{Add} & \ybox{Keep} & \gbox{Drop} \\\midrule
      \code{CannotCallNonFunction} & {0} & {0} & {0} & {17} & {4} & {20} & {1} & {0} & {1} \\
      \code{CannotExtendTable} & {0} & {0} & {0} & {7} & {9} & {10} & {0} & {0} & {1} \\
      \code{CannotInferBinaryOperation} & {0} & {0} & {0} & {1} & {2} & {1} & {7} & {6} & {9} \\
      \code{CountMismatch} & {0} & {0} & {0} & {223} & {48} & {269} & {12} & {2} & {17} \\
      \code{DuplicateTypeDefinition} & {0} & {0} & {0} & {1} & {0} & {0} & {0} & {0} & {0} \\
      \code{ExtraInformation} & {0} & {0} & {0} & {48} & {7} & {39} & {2} & {0} & {1} \\
      \code{FunctionDoesNotTakeSelf} & {0} & {0} & {0} & {4} & {6} & {5} & {0} & {0} & {0} \\
      \code{FunctionExitsWithoutReturning} & {0} & {0} & {0} & {9} & {3} & {6} & {11} & {5} & {8} \\
      \code{GenericError} & {0} & {0} & {0} & {206} & {52} & {203} & {8} & {1} & {9} \\
      \code{IllegalRequire} & {0} & {0} & {0} & {16} & {3} & {33} & {0} & {0} & {2} \\
      \code{IncorrectGenericParameterCount} & {0} & {0} & {0} & {0} & {0} & {0} & {1} & {0} & {1} \\
      \code{MissingProperties} & {0} & {0} & {0} & {11} & {3} & {11} & {6} & {4} & {6} \\
      \code{MissingUnionProperty} & {0} & {0} & {0} & {1} & {0} & {2} & {0} & {0} & {0} \\
      \code{ModuleHasCyclicDependency} & {0} & {0} & {0} & {20} & {9} & {15} & {0} & {0} & {1} \\
      \code{NotATable} & {0} & {0} & {0} & {7} & {2} & {8} & {2} & {0} & {1} \\
      \code{OccursCheckFailed} & {0} & {0} & {0} & {0} & {0} & {0} & {0} & {0} & {1} \\
      \code{OnlyTablesCanHaveMethods} & {0} & {0} & {0} & {1} & {0} & {2} & {0} & {0} & {0} \\
      \code{OptionalValueAccess} & {0} & {0} & {0} & {40} & {52} & {27} & {6} & {2} & {7} \\
      \code{SyntaxError} & {8349} & {314} & {19220} & {983} & {46} & {2658} & {16} & {0} & {71} \\
      \code{TypeMismatch} & {0} & {0} & {0} & {159} & {62} & {129} & {21} & {10} & {28} \\
      \code{TypesAreUnrelated} & {0} & {0} & {0} & {1} & {0} & {0} & {0} & {0} & {1} \\
      \code{UnknownPropButFoundLikeProp} & {0} & {0} & {0} & {26} & {18} & {21} & {1} & {0} & {1} \\
      \code{UnknownProperty} & {0} & {0} & {0} & {347} & {180} & {385} & {19} & {19} & {35} \\
      \code{UnknownRequire} & {0} & {0} & {0} & {91} & {39} & {85} & {12} & {3} & {4} \\
      \code{UnknownSymbol} & {0} & {0} & {0} & {2348} & {496} & {2973} & {49} & {19} & {55}
    \end{tabular}

@@

FILL type error popularity
Mostly syntax errors though!
Estimate \pct{40} are substantial type errors rather than syntax
errors based on the edit range type errors~(\cref{s:type-error-count})
but unfortunately we cannot get an exact number for the overall
or per-module type error counts.

@@

@@

\subsection{Dead Ends: Too Complex}
%% data = out/ctc-info.txt

>>>>

\begin{table}[t]
  \caption{Number of records that add, keep, or drop (remove) specific type errors from their edit range.}
  \label{t:type-error-survival}
  \begin{tabular}{lr@{}r@{}rr@{}r@{}rr@{}r@{}r}
    & \zerowidth{\mnocheck{}} & & & \zerowidth{\mnonstrict{}} & & & \zerowidth{\mstrict{}} & & \\
    & \rbox{Add} & \ybox{Keep} & \gbox{Drop} & \rbox{Add} & \ybox{Keep} & \gbox{Drop} & \rbox{Add} & \ybox{Keep} & \gbox{Drop} \\\midrule
    \code{CannotCallNonFunction} & {0} & {0} & {0} & {17} & {4} & {20} & {1} & {0} & {1} \\
    \code{CannotExtendTable} & {0} & {0} & {0} & {7} & {9} & {10} & {0} & {0} & {1} \\
    \code{CannotInferBinaryOperation} & {0} & {0} & {0} & {1} & {2} & {1} & {7} & {6} & {9} \\
    \code{CountMismatch} & {0} & {0} & {0} & {223} & {48} & {269} & {12} & {2} & {17} \\
    \code{DuplicateTypeDefinition} & {0} & {0} & {0} & {1} & {0} & {0} & {0} & {0} & {0} \\
    \code{ExtraInformation} & {0} & {0} & {0} & {48} & {7} & {39} & {2} & {0} & {1} \\
    \code{FunctionDoesNotTakeSelf} & {0} & {0} & {0} & {4} & {6} & {5} & {0} & {0} & {0} \\
    \code{FunctionExitsWithoutReturning} & {0} & {0} & {0} & {9} & {3} & {6} & {11} & {5} & {8} \\
    \code{GenericError} & {0} & {0} & {0} & {206} & {52} & {203} & {8} & {1} & {9} \\
    \code{IllegalRequire} & {0} & {0} & {0} & {16} & {3} & {33} & {0} & {0} & {2} \\
    \code{IncorrectGenericParameterCount} & {0} & {0} & {0} & {0} & {0} & {0} & {1} & {0} & {1} \\
    \code{MissingProperties} & {0} & {0} & {0} & {11} & {3} & {11} & {6} & {4} & {6} \\
    \code{MissingUnionProperty} & {0} & {0} & {0} & {1} & {0} & {2} & {0} & {0} & {0} \\
    \code{ModuleHasCyclicDependency} & {0} & {0} & {0} & {20} & {9} & {15} & {0} & {0} & {1} \\
    \code{NotATable} & {0} & {0} & {0} & {7} & {2} & {8} & {2} & {0} & {1} \\
    \code{OccursCheckFailed} & {0} & {0} & {0} & {0} & {0} & {0} & {0} & {0} & {1} \\
    \code{OnlyTablesCanHaveMethods} & {0} & {0} & {0} & {1} & {0} & {2} & {0} & {0} & {0} \\
    \code{OptionalValueAccess} & {0} & {0} & {0} & {40} & {52} & {27} & {6} & {2} & {7} \\
    \code{SyntaxError} & {8349} & {314} & {19220} & {983} & {46} & {2658} & {16} & {0} & {71} \\
    \code{TypeMismatch} & {0} & {0} & {0} & {159} & {62} & {129} & {21} & {10} & {28} \\
    \code{TypesAreUnrelated} & {0} & {0} & {0} & {1} & {0} & {0} & {0} & {0} & {1} \\
    \code{UnknownPropButFoundLikeProp} & {0} & {0} & {0} & {26} & {18} & {21} & {1} & {0} & {1} \\
    \code{UnknownProperty} & {0} & {0} & {0} & {347} & {180} & {385} & {19} & {19} & {35} \\
    \code{UnknownRequire} & {0} & {0} & {0} & {91} & {39} & {85} & {12} & {3} & {4} \\
    \code{UnknownSymbol} & {0} & {0} & {0} & {2348} & {496} & {2973} & {49} & {19} & {55}
  \end{tabular}

@@

%% \paragraph{Most Type Errors are Typos}
\begin{itemize}
  \item
    One striking aspect of~\cref{t:type-error-survival} is the large number
    of \code{SyntaxError}s and the related errors \code{UnknownSymbol} and
    \code{UnknownProperty}.
    Most type errors in the data are likely due to typos.
    \Cref{s:type-error-count} discusses this point in detail.
    For now, we remark that \code{SyntaxError} is the only possible error
    in \mnocheck{} mode (by design) and observe that these errors are
    usually in the Drop column and seldom in the Keep column.

  \item
    The numbers are relatively low in general, which means that
    few errors in the dataset overlap with the edit range
    The highest value in the \mstrict{} columns is 71, meaning 71 records
    out of 6,505 total~(\cref{f:dataset-overview} drop a syntax error.
    The \mnonstrict{} and \mnocheck{} columns are similar.
    %% FILL show percentages for all three modes?
    \FILL{} what does this mean? Not sure if bad, or inconclusive.

  \item
    %% Drop >= Add >= Rem ~= 27 triples
    %% other nonzero      ~= 13 triples
    When type errors overlap with the edit range,
    we would hope the number of Adds and Drops is roughly
    equal and the number of Keeps is low.
    This is indeed the case.
    In most Add/Keep/Drop triples, the Drop column has the highest
    value and the Keep column has the lowest value.
    The optimistic conclusion is that creators are fixing the code
    highlighted by errors.

  \item
    After the syntax errors, \code{CountMismatch},
    \code{TypeMismatch}, and \code{GenericError} are the most common
    \mnonstrict{} and \mstrict{} errors.
    \FILL{} so what.

  \item
    \FILL{} which errors break the Drop > Add > Keep hope?

  \item
    (NOTE may want to save relative popularity for next section.)
    Some errors occur more often in \mnonstrict{} than \mstrict{}:
    \code{CannotCallNonFunction},
    \code{CannotExtendTable},
    \code{ExtraInformation},
    \code{FunctionDoesNotTakeSelf},
    \code{IllegalRequire},
    \code{ModuleHasCyclicDependency},
    \code{OptionalValueAccess},
    \code{UnknownPropButFoundLikeProp}.
    \FILL{} is this surprising, or expected from the bigger number of
    \mnonstrict{} records?

  \item
    A few errors are more common in \mstrict{} than \mnonstrict{}:
    \code{CannotInferBinaryOperation},
    \code{FunctionExitsWithoutReturning},
    \code{IncorrectGenericParameterCount}.
    \FILL{} so what.

@@

\FILL{} data seems to say yes.

@@

\FILL{} compare \mnonstrict{} and \mstrict{}.
Can't compare to \mnocheck{} with this table, though, so maybe it make sense to
put this question later.

@@

On a coarser level, \cref{t:type-error-count} reports the percent of
type errors that overlap with the edit range.
For each error type that ever overlaps, the table
shows what percentage it takes relative to all overlaps.

The table does not report \mnocheck{} because all its type errors are
by definition \code{SyntaxError}s.

\begin{itemize}
  \item
    The most popular errors are syntax errors.
    This is undoubtedly due to the stochiastic nature
    of our telemetry, which has no way to avoid
    reporting edits that happen in the middle of things.
    For example, a creator who writes a method call
    \code{bucket.countFish()} letter-by-letter will
    get type analysis to find errors like \code{UnknownSymbol(buck)}
    and \code{UnknownProperty(bucket.c)}.
    These are transient issues, but telemetry may catch them.

  \item
    \FILL{} relative mismatches between NS and S?

  \item
    \FILL{} which errors that NEVER showed up?

@@

\paragraph{Internal Limits, Code Too Complex}
%% data = out/ctc-info.txt

@@
